### 2019-09-25 manuscript.Rmd
update code, gitignore - RSOS df does not have all versions

---
title: "Open Peer Review"
author: "Daniel Lakens"
date: "22-9-2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(dplyr)
library(formattable)
library(tidyr)
library(ggplot2)
library(here)

here::here()

```

## Accessing Open Peer Reviews

PeerJ assigns all articles a number from 1, increasing consequtively with each published manuscript. Reviews are always accessible in html (i.e., at https://peerj.com/articles/1/reviews for the first article). Using text-mining tools in R (cite R, RCurl, XML, stringer) we downloaded all reviews and stored them as individual text files.

```{r}
# Read open science and open biology rds datafiles
TRS_data_os <- readRDS(file = "royal_society_data_os.rds")
TRS_data_ob <- readRDS(file = "royal_society_data_ob.rds")

# Combine both datasets into one.
TRS_data <- rbind(TRS_data_os, TRS_data_ob)

PeerJ_data <- readRDS(file = "peerj_data.rds")

TRS_data_sorted <- TRS_data[ with( TRS_data , order ( df_link)),]

```

Older Royal Society journals have the comment: Note: This manuscript was transferred from another Royal Society journal without peer review.
What does this mean?